# Importing the data

In this chapter, we will import the data using the package IOHAnalyzer and the data available in the repositories below.

* https://github.com/IOHprofiler/IOHdata for the PBO dataset
* https://numbbo.github.io/data-archive/bbob/ for the BBOB 2019 and 2009

The processed data from this chapter is available in the folder data of the repository https://github.com/davidissamattos/evo-irt

```{r message=F}
library(IOHanalyzer)
```

To import this data, we use the function below to convert to csv format for easier analysis.

This gives a summarized values of the number of runs and the number of succeeded runs.

```{r}
import_data <- function(location, folder=F){
  #dsl is a datasetlist from IOHanalyzer
  if(folder==T)
    dsl <- DataSetList(location)
  else
    dsl<- readRDS(location)
  #each dataset is an element of this list, e.g. dsl[[1]]
  #each dataset is a list with three items RT (runtime value for fixed target) FV (function value for fixed budget) and PAR (parameters)
  ds_list_fv <- lapply(dsl,get_FV_overview)
  ds_list_rt <- lapply(dsl,get_RT_overview)
  fv <- do.call(rbind,ds_list_fv)
  rt <- do.call(rbind,ds_list_rt)
  out <- list(rt=rt,
             fv=fv)
  }
```

## PBO dataset

The IOHData repository already has some benchmarks in the RDS format which we will use here.

```{r eval=F}
d_pbo <- import_data('data/IOHdata-master/iohprofiler/2019gecco-ins11-1run.rds', folder = F)
d_pbo_fv <- d_pbo$fv
d_pbo_fv$succ <- d_pbo_fv$succ %>% 
  replace_na(0)
write_csv(d_pbo_fv,'data/pbo.csv')
```

## BBOB 2019

Import folder
```{r eval=F}
d_bbob1 <- import_data('data/BBOB/Adaptive-Two-Mode_bbob_Bodner/', folder = T)
d_bbob2 <- import_data('data/BBOB/adapt-Nelder-Mead-scipy-2019_bbob_Varelas_Dahito/', folder = T)
d_bbob3 <- import_data('data/BBOB/BFGS-scipy-2019_bbob_Varelas_Dahito/minimize_on_bbob_budget100000xD/', folder = T)
d_bbob4 <- import_data('data/BBOB/CG-scipy-2019_bbob_Varelas_Dahito/', folder = T)
d_bbob5 <- import_data('data/BBOB/COBYLA-scipy-2019_bbob_Varelas_Dahito/', folder = T)
d_bbob6 <- import_data('data/BBOB/DE-scipy-2019_bbob_Varelas_Dahito/', folder = T)
d_bbob7 <- import_data('data/BBOB/L-BFGS-B-scipy-2019_bbob_Varelas_Dahito/', folder = T)
d_bbob8 <- import_data('data/BBOB/Nelder-Mead-scipy-2019_bbob_Varelas_Dahito/', folder = T)
d_bbob9 <- import_data('data/BBOB/Powell-scipy-2019_bbob_Varelas_Dahito/', folder = T)
d_bbob10 <- import_data('data/BBOB/RS-4-initIn0_bbob_Brockhoff_Hansen/', folder = T)
d_bbob11 <- import_data('data/BBOB/RS-5-initIn0_bbob_Brockhoff_Hansen/', folder = T)
d_bbob12 <- import_data('data/BBOB/RS-6-initIn0_bbob_Brockhoff_Hansen/', folder = T)
d_bbob13 <- import_data('data/BBOB/TNC-scipy-2019_bbob_Varelas_Dahito/', folder = T)
```

Now we merge these datasets into a single one
```{r eval=F}
d_bbob <- rbind(d_bbob1$fv,
                d_bbob2$fv,
                d_bbob3$fv,
                d_bbob4$fv,
                d_bbob5$fv,
                d_bbob6$fv,
                d_bbob7$fv,
                d_bbob8$fv,
                d_bbob9$fv,
                d_bbob10$fv,
                d_bbob11$fv,
                d_bbob12$fv,
                d_bbob13$fv
                )
#NA in succ will become 0
d_bbob$succ <- d_bbob$succ %>% 
  replace_na(0)
write_csv(d_bbob,'data/bbob2019.csv')
```

## BBOB 2009

```{r eval=F}
bbob_2009_dirs <- list.dirs(path = "data/BBOB2009", full.names = TRUE, recursive = F)
bbob2009<-NULL
for(i in seq(1:length(bbob_2009_dirs))){
  d <- import_data(bbob_2009_dirs[i], folder = T)
  bbob2009 <- rbind(bbob2009,d$fv)
}
bbob2009$succ <- bbob2009$succ %>% 
  replace_na(0)

write_csv(bbob2009,'data/bbob2009.csv')
```

