# IRT Stan

## 2PL model

The description and interpretation of the model is available in the paper. Here we are just generating the plots in the paper

Function to generate the item characteristic curve. 
```{r}
icc_2pl <- function(a, b, label, thetamin=-5, thetamax=5, step=0.1){
  x<-seq(from=thetamin, to=thetamax, by=step)
  y<-exp(a*(x-b))/(1+exp(a*(x-b)))
 out<-data.frame(
   Prob=y,
   Ability=x,
   label=rep(label,length(x))) 
 return(out)
}
```

### Impact of item difficulty

Let's plot three item characteristic curves with different difficulties and the same discrimination.

First, we generate the curves:
```{r}
icc_easy <- icc_2pl(a=3,b=-3, label = 'Easy (b=-3)')
icc_medium <- icc_2pl(a=3,b=0, label = 'Medium (b=0)')
icc_hard <- icc_2pl(a=3,b=3, label = 'Hard (b=3)')
icc_b <- rbind(icc_easy,icc_medium,icc_hard)
```

Then we plot it
```{r}
ggplot(icc_b, aes(x=Ability, y=Prob, color=label))+
  geom_line()+
  labs(
    title='Item Characteristic Curve',
    # subtitle = 'Impact of difficulty parameter',
    x=unname(TeX("Ability ($\\theta$)")),
    y='Probability of solving',
    color = 'Difficulty'
  )+
  theme_bw()+
  theme(legend.position = 'bottom')
```

### Impact of item discrimination

Let's plot three item characteristic curves with different discrimination and the same difficulty

First, we generate the curves:
```{r}
icc_lowdisc <- icc_2pl(a=0.3,b=0, label = 'Low (a=0.3)')
icc_mediumdisc <- icc_2pl(a=1,b=0, label = 'Medium (a=1)')
icc_highdisc <- icc_2pl(a=3,b=0, label = 'High (a=3)')
icc_a <- rbind(icc_lowdisc,icc_mediumdisc,icc_highdisc)
```

Then we plot them:
```{r}
ggplot(icc_a, aes(x=Ability, y=Prob, color=label))+
  geom_line()+
  labs(
    title='Item Characteristic Curve',
    x=unname(TeX("Ability ($\\theta$)")),
    y='Probability of solving',
    color = 'Discrimination'
  )+
  theme_bw()+
  theme(legend.position = 'bottom')
```

### Item Information

We can calculate the item information for the 2PL. We use the direct formula for the information from the paper.

```{r}
item_info <- function(a,b,thetamin=-5, thetamax=5,step=0.1){
  theta <- seq(from=thetamin, to=thetamax, by=step)
  p<-exp(a*(theta-b))/(1+exp(a*(theta-b)))
  q<-1-p
  info <- a^2*p*q
  return(list(info=info,theta=theta))
}
```

Plotting this for a few different item characteristic curves (different values of a and b).

First we create a long data frame with one column for the information
```{r}
d_item_info_pars <- data.frame(a=c(0.3,1,1.2,0.8,0.6),
                          b=c(1.5,-2.2,1,-0.4,0.9))
d_item_info<-NULL
for(row in 1:nrow(d_item_info_pars)){
  info_list <- item_info(d_item_info_pars$a[row],d_item_info_pars$b[row])
  d_item_info<-rbind(
    d_item_info,
    data.frame(
      information=info_list$info, 
      theta=info_list$theta,
      label=rep(
        glue::glue('a=',d_item_info_pars$a[row],
                   ',b=',d_item_info_pars$b[row]),
        length(info_list$info)
        )
      )
    )
}
```

Now we can plot:
```{r}
ggplot(d_item_info, aes(x=theta, y=information, color=label))+
  geom_line()+
  labs(
    title='Item Information Curve',
    x=unname(TeX("Ability ($\\theta$)")),
    y='Information',
    color = 'Item'
  )+
  theme_bw()+
  theme(legend.position = 'bottom')+
  guides(color=guide_legend(nrow=2,byrow=TRUE))
```

### Test information 

For a test, we can calculate the test information as the sum of the individual items.

First let's pivot wider so we can sum the columns:
```{r}
d_item_info_wider <- d_item_info %>%
  pivot_wider(names_from = 'label', values_from = 'information') %>%
  mutate(TestInfo = dplyr::select(. ,starts_with('a=')) %>% rowSums())
```

Now we can plot the test information curve:
```{r}
ggplot(d_item_info_wider, aes(x=theta, y=TestInfo)) +
  geom_line(aes(y=TestInfo))+
    labs(
    title='Test Information Curve',
    x=unname(TeX("Ability ($\\theta$)")),
    y='Information'
  )+
  theme_bw()
```


## Bayesian 2PL with multiple attempts

The Bayesian two parameter logistic item response theory model with multiple attempts is represented by the following equations.

$$
y_{i,p} \sim \text{Binomial}(N_{i,p},\mu_{i,p}) \\
\mu_{i,p} = \dfrac{\exp(a_i*(b_i-\theta_p))}{1+ \exp(a_i*(b_i-\theta_p))} \\
a_i \sim \text{Half-}\mathcal{N}(0,3) \\
b_i \sim \mathcal{N}(0,3)\\
\theta_p \sim \mathcal{N}(0,3)
$$

The Stan code for this model is shown below:

```{r echo=F}
#Show stan code from separate file in this block
show_stan_code <- function(filename)
{
  sourcecode <- paste(readLines(filename), collapse="\n")
  cat(sourcecode)
}
```

```{r}
show_stan_code('models/irt2pl.stan')
```

