# Case Study I: BBOB 2019

## Importing the data

To illustrate and make the analysis, we will the number of dimensions equal to 5 (since the benchmark functions are all scalable). To do an analysis with different dimensions just change the code here

```{r message=F}
d_bbob <- read_csv('data/bbob2019.csv') %>% 
  select(algId, DIM, funcId, runs, succ, budget) %>% 
  filter(DIM==5) %>%  
  mutate(algId_index = as.integer(as.factor(algId)))

#vector with the names in order
benchmarks <- seq(1,24)
algorithms <- levels(as.factor(d_bbob$algId))
```

## Preparing the Stan data

Creating a list for Stan.
```{r}
bbob_standata <- list(
  N = nrow(d_bbob),
  y_succ = as.integer(d_bbob$succ),
  N_tries = as.integer(d_bbob$runs),
  p = d_bbob$algId_index,
  Np = as.integer(length(unique(d_bbob$algId_index))),
  item = as.integer(d_bbob$funcId),
  Nitem = as.integer(length(unique(d_bbob$funcId)))
)
```

Calling the model with cmdstanr

```{r eval=F}
irt2pl <- cmdstan_model('models/irt2pl.stan') 
fit_bbob <- irt2pl$sample(
  data= bbob_standata,
  chains = 4,
  iter_sampling = 4000,
  parallel_chains = 4,
  max_treedepth = 15
)
fit_bbob$save_object(file='fitted/bbob5.RDS')
```

To load the fitted model (to save time in compiling this document)

```{r}
fit_bbob<-readRDS('fitted/bbob5.RDS')
```

## Diagnostics

Getting the draws from the posterior

```{r}
draws_a <- fit_bbob$draws('a')
draws_b <- fit_bbob$draws('b')
draws_theta <- fit_bbob$draws('theta')
```

### Traceplots

```{r}
mcmc_trace(draws_a)
```

```{r}
mcmc_trace(draws_b)
```

```{r}
mcmc_trace(draws_theta)
```

### Rhat and Effective samples

```{r}
fit_bbob$summary() %>% 
  kable(caption='Summary values fit of the model, including effective samples and Rhat', 
      booktabs=T,
      digits =3,
      format='html') %>% 
  kable_styling() %>% 
  scroll_box()
```


## Results

Let's get some summary descriptive statistics of the posterior
```{r}
fit_summary_a_b <- fit_bbob$summary(c('a','b'))
fit_summary_a <- fit_bbob$summary(c('a'))
fit_summary_b <- fit_bbob$summary(c('b'))
fit_summary_theta <- fit_bbob$summary(c('theta'))
```


### Difficulty and discrimination

Table for the benchmark functions

```{r}
table_benchmarks <- fit_summary_a_b %>% 
  select('Benchmark ID'=variable, 
         Median=median,
         'CI 5%'=q5,
         'CI 95%'=q95)

table_benchmarks$'Benchmark ID'<-rep(benchmarks,2)

kable(table_benchmarks,
      caption='Summary values of the discrimination and difficulty level parameters for the BBOB benchmarks', 
      booktabs=T,
      digits =3,
      format='html',
      linesep = "") %>% 
  kable_styling() %>% 
  pack_rows("Discrimination value (a)",1,24) %>% 
  pack_rows("Difficulty level (b)",25,48)
```

A more visual representation.
```{r}
mcmc_intervals(draws_a) +
  scale_y_discrete(labels=benchmarks)+
  labs(x='Discrimination parameter (a)',
       y='Benchmark function ID',
       title='Discrimination parameter distribution (BBOB)')
```

```{r}
mcmc_intervals(draws_b) +
  scale_y_discrete(labels=benchmarks)+
  labs(x='Difficulty level parameter (b)',
       y='Benchmark function ID',
       title='Difficulty level parameter distribution (BBOB)')
```


### Ability

Creating a table:

```{r}
table_algorithms <- fit_summary_theta %>% 
  select(Algorithms=variable, 
         Median=median,
         'CI 5%'=q5,
         'CI 95%'=q95)

table_algorithms$Algorithms <- algorithms

kable(table_algorithms,
      caption='Summary values of the ability level of the algorithms (BBOB)',
      booktabs=T,
      digits =3,
      format='html',
      linesep = "") %>% 
  kable_styling() 
```

A more visual representation.
```{r}
mcmc_intervals(draws_theta) +
  scale_y_discrete(labels=algorithms)+
  labs(x=unname(TeX("Ability level ($\\theta$)")),
       y='Algorithm',
       title='Ability level parameter distribution (BBOB)')
```

### Item information 

First let's create a few helper functions to calculate the item information

```{r}
p_info <- function(a,b, theta){
     return(exp(a*(theta-b))/(1+exp(a*(theta-b))))
  }
q_info <- function(a,b, theta){
    return(1-p_info(a,b, theta))
  }
#a and b are a vector of 3 a[1] is lower q05 a[2] is median and a[3] is q95
#return a data frame ready to be plottted
item_info_with_intervals <- function(a,b,item, thetamin=-5, thetamax=5,step=0.1){
  theta <- seq(from=thetamin, to=thetamax, by=step)
  info_median <- a[1]^2*p_info(a[1],b[1],theta)*q_info(a[1],b[1],theta)
  info_lower <- a[2]^2*p_info(a[2],b[2],theta)*q_info(a[2],b[2],theta)
  info_higher <- a[3]^2*p_info(a[3],b[3],theta)*q_info(a[3],b[3],theta)
  
  out<- data.frame(Information= c(info_lower,info_median,info_higher),
                   theta=c(theta,theta,theta),
                   pars=c(rep('q05',length(theta)),
                           rep('median',length(theta)),
                           rep('q95',length(theta))),
                   item=c(rep(item,length(theta)),
                          rep(item,length(theta)),
                          rep(item,length(theta))))
  return(out)
}
```

Creating a single data frame
```{r}
item_information_df <- NULL
for(i in seq(1:length(benchmarks))){
  a<-as.matrix(fit_summary_a[i,c(3,6,7)])
  b<-as.matrix(fit_summary_b[i,c(3,6,7)])
  iinfo <- item_info_with_intervals(a=a,b=b,item = i,thetamin = -7, thetamax = 5)
  item_information_df <- rbind(item_information_df,iinfo)
}

```

Now we can create an information plot for every item

```{r}
item_information_df %>% 
  pivot_wider(names_from = 'pars', values_from = 'Information') %>% 
  ggplot(aes(x=theta))+
    geom_line(aes(y=median), color='black')+
    facet_wrap(~item,
               ncol=4) +
    labs(title='Item information curve (BBOB)',
         x=unname(TeX("Ability ($\\theta$)")),
         y='Information',
         color='Information interval')+
    theme_bw() +
    theme(legend.position = 'bottom')
```

### Test information

We can also look at the test information. First, we need to pivot wider so we can sum the items

```{r}
test_information_df <- item_information_df %>% 
  pivot_wider(names_from = 'item', values_from = 'Information') %>% 
  mutate(TestInfo = dplyr::select(., -theta, -pars) %>% rowSums()) %>% 
  dplyr::select(theta, pars, TestInfo)
```

Now that we have calculated the test parameters we can plot the test information

First let's get a horizontal line to show where the algorithms median ability lies
```{r}
alg_median <- fit_summary_theta %>% 
  mutate(Algorithm=algorithms) %>% 
  select(Algorithm, median) 
```


```{r}
test_information_df %>% 
  dplyr::select(theta, pars, TestInfo) %>% 
  pivot_wider(names_from = 'pars', values_from = 'TestInfo') %>% 
  ggplot(aes(x=theta)) +
  geom_line(aes(y=median))+
  geom_vline(data=alg_median, aes(xintercept=median,color=Algorithm),linetype='dashed')+
  labs(
    title='Test Information Curve (BBOB)',
    x=unname(TeX("Ability ($\\theta$)")),
    y='Test information',
    color='Algorithm median'
  )+
  theme_bw()+
  guides(color=guide_legend(nrow=5,byrow=TRUE))+
  theme(legend.position = 'bottom')
```
