--- 
title: 'Online appendix for the paper: "On the Assessment of Benchmark Suites for Algorithm Comparison"'
author: "David Issa Mattos, Lucas Ruud, Jan Bosch and Helena Holmström Olsson."
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: book
output:
  bookdown::gitbook: default
bibliography: [book.bib]
biblio-style: apalike
link-citations: yes
description: 'This is the online appendix for the paper: "On the Assessment of Benchmark Suites for Algorithm Comparison"'
always_allow_html: true
---

```{r include=FALSE,message=F}
# automatically create a bib database for R packages
knitr::opts_chunk$set(cache=TRUE)
library(tidyverse)
library(knitr)
library(posterior)
library(bayesplot)
library(kableExtra)
library(latex2exp)
library(cmdstanr)
seed<-123
set.seed(seed)
```

# Foreword

This is the online appendix for the paper *"On the Assessment of Benchmark Suites for Algorithm Comparison"*. It contains a commented and reproducible code for all the analysis, tables and plots presented in the paper and additional content.

The code used to generate this appendix is available at the repository: https://github.com/davidissamattos/evo-irt

## Session info

This appendix is compiled automatically and the following session information was used to generate this appendix:

```{r}
sessionInfo()
```




